---
title: "EDA work_1"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    embed-resources: true
execute:
  include: true
  eval: true    
  warning: false
  message: false
---

# Intro

This file is to review part of the data provided by Swire to 

# Libraries and intial data loading

## Libraries
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(skimr)
library(janitor)
```

## Data Loading

Here are the data sets I'll be reviewing in this EDA workbook

### Material data and overview
```{r data loaded and copied}
o_material <- read_csv("~/Masters/Fall 2025/IS6813 Capstone 3/data/material.csv")
material <- o_material #copied

summary(material)
```

Distinct counting 
```{r material distinct count}
#unique count of material_id
n_distinct(material$MATERIAL_ID)
```

All objects listed in material are distinct. 

Updating the MATERIAL_ID to numeric and factoring other columns for analysis
```{r}
#updating column to numeric
material <- material %>%
  mutate(MATERIAL_ID = as.numeric(MATERIAL_ID))

#Factoring other columns for analysis
material <- material %>%
  mutate(across(c(PACK_TYPE_DESC, TRADE_MARK_DESC, FLAVOUR_DESC, PACK_SIZE_DESC, BEV_CAT_DESC), 
                as.factor))
```

#### Missing data
```{r material missing data}
# Filter rows with any NA, show all columns
material %>%
  filter(if_any(everything(), is.na)) %>%
  arrange(PACK_TYPE_DESC, TRADE_MARK_DESC, PACK_TYPE_DESC) 

# results has 134 missing, only in Bev_CAT_ DESC. the trade_mark_desc and flavor_desc only has 1 variable for each column Sunny Sip and Passionfruit (respectfully) that match the NA
```


#### Small analysis

Reviewing the common combinations, not considering the MATERIAL_ID, as this is unique for all of the data in the set
```{r material common combo}
# Grouping for most common combinations in the data set, not considering the material_id
mat_combo_count <- material %>%
  count(PACK_TYPE_DESC, TRADE_MARK_DESC, FLAVOUR_DESC, PACK_SIZE_DESC, BEV_CAT_DESC)

# arranging the order to most common duplicates (n > 1)
material %>%
  count(PACK_TYPE_DESC, TRADE_MARK_DESC, FLAVOUR_DESC, PACK_SIZE_DESC, BEV_CAT_DESC) %>%
  filter(n >1 ) %>%
  arrange(desc(n))
```


### Google Analytics data
```{r google analytics data loading and summary}
o_data_google <- read_csv("~/Masters/Fall 2025/IS6813 Capstone 3/data/google_analytics.csv")

google_analyc <- o_data_google

summary(google_analyc)
```




### Orders data
```{r google analytics data loading and summary}
o_data_google <- read_csv("~/Masters/Fall 2025/IS6813 Capstone 3/data/google_analytics.csv")

google_analyc <- o_data_google

summary(google_analyc)
```

### Sales data